---
title: "Assignment 1"
author: "Adnan Selak"
date: "November 13, 2015"
output: html_document
---
## Loading and preprocessing Data

```{r load libraries, echo=TRUE}
library(dplyr)
library(ggplot2)
library(scales)
library(reshape2)
library(cowplot)
```
```{r preprocess data, echo=TRUE}
setwd("C:\\Users\\Selak\\Dropbox\\Coursera\\04 - Reproducible Research\\project_1")
raw_data <- read.csv("activity.csv", header = TRUE)
raw_data$date <- as.Date(raw_data$date, format = "%Y-%m-%d")
## Total Means...
sum_data <- summarise(group_by(raw_data,date),sum_steps = sum(steps, na.rm = TRUE))
```

## Total Number of Steps Per Day
NA values are treated as 0.

```{r plot sums, fig.height=3, echo=TRUE}

ggplot(sum_data) + geom_histogram(aes(x=date, y=sum_steps), stat="identity") + xlab("Date") + ylab("Total Number of Steps") + scale_x_date()
```

```{r mean and median, echo=TRUE}
mean_data <- mean(sum_data$sum_steps)
median_data <- median(sum_data$sum_steps)
```
Total Number of Steps daily mean is `r format(mean_data,digits=6,nsmall=2)` and median is `r format(median_data,digits=6,nsmall=2)`.

## Daily Activity Pattern
```{r daily activity, fig.height=3, echo=TRUE}
interval_data <- group_by(raw_data,interval)
sum_int_data <- summarise(interval_data,int_mean = mean(steps, na.rm = TRUE), int_median = median(steps,na.rm=TRUE))
p2 <- ggplot(sum_int_data)
p2 <- p2 + geom_line(aes(x=interval, y=int_mean, color = "int_mean")) + geom_line(aes(x=interval, y=int_median, color = "int_median"))
p2 <- p2 + labs(color="Variables") + xlab("5-min Interval") + ylab("Number of Steps")
p2
#Find maximum value and corresponding row index
max_val <- max(sum_int_data$int_mean)
max_int <- which.max(sum_int_data$int_mean)
```
The interval containing the maximum number of steps (average) is `r sum_int_data$interval[max_int]`.

# Data with NAs being replaced.
Total number of records containing NA is `r count(raw_data[rowSums(is.na(raw_data)) > 0,])`.
```{r NA replace an plot, echo=TRUE}
# Replace NAs in steps column with Interval Means for number of steps
merged_data <- merge(raw_data, sum_int_data, by.x = "interval", by.y = "interval")
merged_data$steps[is.na(merged_data$steps)] <- merged_data$int_mean[is.na(merged_data$steps)]
clean_data <- merged_data[,1:3]

## Total Means...
sum_clean_data <- summarise(group_by(clean_data,date),sum_steps = sum(steps, na.rm = TRUE))
## Total Number of Steps Per Day
```
## Total Number of Steps Per Day
```{r plot sums NA, fig.height=3, echo=TRUE}
ggplot(sum_clean_data) + geom_histogram(aes(x=date, y=sum_steps), stat="identity") + xlab("Date") + ylab("Total Number of Steps") + scale_x_date()
```
```{r mean and median NA, echo=TRUE}
mean_data <- mean(sum_clean_data$sum_steps)
median_data <- median(sum_clean_data$sum_steps)
```
Total Number of Steps daily mean is `r format(mean_data,digits=6,nsmall=2)` and median is `r format(median_data,digits=6,nsmall=2)`.

## Weekday versus Weekend Activity Pattern
```{r plot interval NA, fig.height = 3, fig.width=12, echo=TRUE}
# Create DayType variable that holds Weekend or Weekday value based on weekdays function
clean_data <- mutate(clean_data, daytype = ifelse(weekdays(clean_data$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
sum_clean_int_data <- summarise(group_by(clean_data,interval,daytype) ,int_mean = mean(steps, na.rm = TRUE), int_median = median(steps,na.rm=TRUE))
weekend_data <- filter(sum_clean_int_data, daytype == "Weekend")
weekday_data <- filter(sum_clean_int_data, daytype == "Weekday")
p1 <- ggplot() + geom_line(data = weekday_data, aes(x = interval, y = int_mean, color = "Mean" )) + 
    geom_line(data = weekday_data, aes(x = interval, y = int_median, color = "Median" ))
p1 <- p1 + labs(color="Variables") + xlab("5-min Interval") + ylab("Number of Steps")
p2 <- ggplot() + geom_line(data = weekend_data, aes( x=interval, y=int_mean, color = "Mean" )) + 
    geom_line(data = weekend_data, aes( x=interval, y=int_median, color = "Median"))
p2 <- p2 + labs(color="Variables") + xlab("5-min Interval") + ylab("Number of Steps")

plot_grid(p1,p2,labels=c("Weekday","Weekend"),ncol = 2, nrow = 1)
```
